---
title: "Interlab study"
subtitle: "Test run"
author: "Hugo Pineda Hernandez"
date: "`r strftime(Sys.Date(), '%B %d %Y')`"
output: 
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning=FALSE, message = FALSE
  
)

library("tidyverse")
library(ggthemes)
library(cowplot)
library(ggridges)
theme_set(theme_light(14))
```

```{r load_data, include=FALSE}
# Load project settings
source(here::here("settings.R"))
# load functions
source(file.path(data.raw.dir, "experiments.R"))

df.pr.od.raw <- load_plate_reader_od()

df.pr.fl.raw <- load_plate_reader_fl()

df.spectrophotometer <- load_spectrophotometer() 

df.full.spectrum.raw <- load_full_spectrum()
```

```{r}

df.pr.od.blank <- df.pr.od.raw %>% 
  filter(!is.na(OD_730), sample_type == "blank") %>% 
  group_by(experiment_date, location, time_h) %>% 
  summarise(
    blank_OD_730 = mean(OD_730)
  ) %>% 
  ungroup()
  
df.pr.od <- df.pr.od.raw %>% 
  filter(!is.na(OD_730), sample_type == "sample") %>% 
  left_join(df.pr.od.blank, by = c("experiment_date", "location", "time_h")) %>% 
  mutate(
    # correct with blank
    OD_730_bc = OD_730 - blank_OD_730
  ) 
  
```

```{r}

df.pr.fl.blank <- df.pr.fl.raw %>% 
  filter(sample_type == "blank") %>% 
  group_by(experiment_date, location, time_h) %>% 
  summarise(
    blank_fl = mean(fl)
  ) %>% 
  ungroup()
  

df.pr.fl <- df.pr.fl.raw %>% 
  filter(sample_type == "sample") %>% 
   left_join(df.pr.fl.blank,
              by = c("experiment_date", "location", "time_h")) %>% 
  mutate(
    # correct with blank
    fl_bc = fl - blank_fl
  ) 
```


```{r}
df.pr <- full_join(
  df.pr.od, 
  df.pr.fl,
  by = c("strain", "induction", "bio_replicate", "location", "tech_replicate", "experiment_date", "experiment_id", "time_h", "sample_type")
  ) %>% 
    mutate(
    fl_od = fl_bc / OD_730_bc
  ) 
```



```{r}

color.scheme.4 <- c("Amsterdam" = "#D55E00", 
                  "Düsseldorf" = "#CC79A7",
                  "Jena" = "#009E73",
                  "Seville" = "#0072B2")

color.scheme.3 <- c("Amsterdam" = "#D55E00", 
                  "Jena" = "#009E73",
                  "Seville" = "#0072B2")


plot_summary_location_strain <- function(){
  list(
  stat_summary(fun.data = "mean_cl_boot", geom = "point", alpha = 0.6),
  stat_summary(fun.data = "mean_sdl", geom = "errorbar", alpha = 0.6),
  stat_summary(fun.data = "mean_cl_boot", geom = "line", alpha = 0.6),
  facet_grid(location~strain, scales = "free_y"),
    scale_colour_colorblind(),
  theme(legend.position = "bottom")
  )
}

plot_summary_ind_strain <- function(colors.scheme){
  list(
  stat_summary(fun.data = "mean_cl_boot", geom = "point", alpha = 0.6),
  stat_summary(fun.data = "mean_sdl", geom = "errorbar", alpha = 0.6),
  stat_summary(fun.data = "mean_cl_boot", geom = "line", alpha = 0.6),
  facet_grid(induction~strain),
  expand_limits(y = 0),
  scale_colour_manual(
    breaks = names(colors.scheme),
    values = colors.scheme
    ),
  theme(legend.position = "bottom")
  )
}
```


# Plate reader data

## Plate reader raw data

### OD<sub>730</sub>

```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr.od %>% 
  ggplot(aes(x = time_h, y = OD_730, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = p.label.od.730
  )
```

```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr.od %>% 
  ggplot(aes(x = time_h, y = OD_730, colour = location)) +
 plot_summary_ind_strain(colors.scheme = color.scheme.3) + 
  labs(
    x = p.label.time.h,
    y = p.label.od.730,
    colour = ""
  ) 
```

### Fluorescence

```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr.fl %>% 
  ggplot(aes(x = time_h, y = fl, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = "Raw flourescence"
  ) 
```


```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr.fl %>% 
  ggplot(aes(x = time_h, y = fl, colour = location)) +
   plot_summary_ind_strain(colors.scheme = color.scheme.4) + 
  labs(
    x = p.label.time.h,
    y = "Raw flourescence",
    colour = ""
  ) 
```


## Plate reader normalized data


### OD<sub>730</sub> corrected by blank

```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr %>% 
  filter(!is.na(OD_730_bc)) %>% 
  ggplot(aes(x = time_h, y = OD_730_bc, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = p.label.od.730
  ) 
```


```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr %>% 
  filter(!is.na(OD_730_bc)) %>%
  ggplot(aes(x = time_h, y = OD_730_bc, colour = location)) +
   plot_summary_ind_strain(colors.scheme = color.scheme.3) + 
  labs(
    x = p.label.time.h,
    y = p.label.od.730,
    colour = ""
  ) 
```


### Background corrected flourescence

```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr %>% 
  ggplot(aes(x = time_h, y = fl_bc, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = "Background corrected flourescence"
  ) 
```


```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr %>% 
  ggplot(aes(x = time_h, y = fl_bc, colour = location)) +
   plot_summary_ind_strain(colors.scheme = color.scheme.4) + 
  labs(
    x = p.label.time.h,
    y = "Background corrected flourescence",
    colour = ""
  ) 
```



## Background corrected flourescence and normalized by OD<sub>730</sub>
 
 
```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr %>% 
  filter(!is.na(fl_od)) %>%
  ggplot(aes(x = time_h, y = fl_od, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = p.label.fl.od
  ) 
```


```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr %>% 
   filter(!is.na(fl_od)) %>%
  ggplot(aes(x = time_h, y = fl_od, colour = location)) +
   plot_summary_ind_strain(colors.scheme = color.scheme.3) + 
  labs(
    x = p.label.time.h,
    y = p.label.fl.od,
    colour = ""
  ) 
```



## Normalization by PJ23100 -

```{r}
df.pr.sum <- df.pr %>% 
  group_by(experiment_date, location, time_h, strain, induction) %>% 
  summarise(
    fl_od = mean(fl_od)
  )
```

```{r}
df.pr.sum.j23 <- df.pr.sum %>% 
  ungroup() %>% 
  filter(strain == "J23100", induction == "-") %>% 
  rename(fl_od_j23 = fl_od) %>% 
  select(-c(strain, induction))
```

```{r}
df.pr.norm.j23 <- left_join(
  df.pr.sum,
  df.pr.sum.j23, 
  by = c("experiment_date", "location", "time_h")
) %>% 
  mutate(
    fl_od_j23_norm = fl_od /  fl_od_j23
  )
```



```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr.norm.j23 %>% 
  filter(!is.na(fl_od_j23_norm)) %>%
  ggplot(aes(x = time_h, y = fl_od_j23_norm, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = p.label.norm.fl.od
  ) 
```


```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr.norm.j23 %>% 
  filter(!is.na(fl_od_j23_norm)) %>%
  ggplot(aes(x = time_h, y = fl_od_j23_norm, colour = location)) +
   plot_summary_ind_strain(colors.scheme = color.scheme.3) + 
  labs(
    x = p.label.time.h,
    y = p.label.norm.fl.od,
    colour = ""
  ) 
```


## Normalization by PJ23100 - and t0

```{r}
df.pr.t0 <- df.pr.norm.j23 %>% 
  ungroup() %>% 
  filter(time_h == 0) %>% 
  select(
    fl_od_t0 = fl_od_j23_norm,
    experiment_date, 
    location,
    strain,
    induction
    
  )
```

```{r}
df.pr.norm.j23.t0 <- df.pr.norm.j23 %>% 
  left_join(df.pr.t0) %>% 
  mutate(
    fl_od_j23_t0_norm = fl_od_j23_norm / fl_od_t0
  )
```




```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr.norm.j23.t0 %>% 
  filter(!is.na(fl_od_j23_norm)) %>%
  ggplot(aes(x = time_h, y = fl_od_j23_t0_norm, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = p.label.norm.fl.od
  ) 
```


```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr.norm.j23.t0 %>% 
  filter(!is.na(fl_od_j23_norm)) %>%
  ggplot(aes(x = time_h, y = fl_od_j23_t0_norm, colour = location)) +
   plot_summary_ind_strain(colors.scheme = color.scheme.3) + 
  labs(
    x = p.label.time.h,
    y = p.label.norm.fl.od,
    colour = ""
  ) 
```


## Normalization by t0


```{r}
df.pr.t0 <-  df.pr %>% 
  ungroup() %>% 
  filter(time_h == 0) %>% 
  select(
    fl_od_t0 = fl_od,
    experiment_date, 
    location,
    strain,
    tech_replicate,
    induction
  )
```

```{r}
df.pr.t0.norm <- df.pr %>% 
  left_join(df.pr.t0, 
            by = c("strain", "induction", 
                   "location", "tech_replicate", 
                   "experiment_date")) %>% 
  mutate(
    fl_od_t0_norm = fl_od / fl_od_t0
  )
```



```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr.t0.norm %>% 
  filter(!is.na(fl_od_t0_norm)) %>%
  ggplot(aes(x = time_h, y = fl_od_t0_norm, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = p.label.norm.fl.od
  ) 
```


```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr.t0.norm %>% 
  filter(!is.na(fl_od_t0_norm)) %>%
  ggplot(aes(x = time_h, y = fl_od_t0_norm, colour = location)) +
   plot_summary_ind_strain(colors.scheme = color.scheme.3) + 
  labs(
    x = p.label.time.h,
    y = p.label.norm.fl.od,
    colour = ""
  ) 
```


### Removing EVC from figures


```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr.t0.norm %>% 
  filter(!is.na(fl_od_t0_norm), strain != "EVC") %>%
  ggplot(aes(x = time_h, y = fl_od_t0_norm, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = p.label.norm.fl.od
  ) 
```


```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr.t0.norm %>% 
  filter(!is.na(fl_od_t0_norm), strain != "EVC") %>%
  ggplot(aes(x = time_h, y = fl_od_t0_norm, colour = location)) +
   plot_summary_ind_strain(colors.scheme = color.scheme.3) + 
  labs(
    x = p.label.time.h,
    y = p.label.norm.fl.od,
    colour = ""
  ) 
```

# Spectrophotometer OD<sub>730</sub>

```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.spectrophotometer %>% 
  filter(!is.na(OD_730)) %>% 
  ggplot(aes(x = time_h, y = OD_730, colour = induction)) +
  plot_summary_location_strain() + 
  labs(
    x = p.label.time.h,
    y = p.label.od.730
  )
```

```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.spectrophotometer %>% 
  filter(!is.na(OD_730)) %>% 
  ggplot(aes(x = time_h, y = OD_730, colour = location)) +
  plot_summary_ind_strain(colors.scheme = color.scheme.3) + 
  labs(
    x = p.label.time.h,
    y = p.label.od.730
  )
```

# Full spectrum

## Normalized each measurement from 0 to 1

```{r}
df.full.spectrum <- df.full.spectrum.raw %>% 
  filter(!is.na(abs)) %>% 
  group_by(location, experiment_date, strain, induction, time_h) %>% 
  mutate(
    norm_abs = (abs  - min(abs)) / (max(abs) - min(abs))
   ) %>% 
  ungroup() %>% 
  mutate(
    location = stringr::str_to_title(location),
    induction = ifelse(is.na(induction), "Time 0", induction)
  )

```

### Time 0

```{r, fig.width=8.4375, fig.height=7.916667, out.width="100%"}
df.full.spectrum %>% 
  filter(time_h == 0) %>% 
  ggplot(aes(x = nm, y = norm_abs, colour = location)) +
  geom_line() +
  facet_wrap(~strain) +
  scale_colour_manual(
    breaks = names(color.scheme.3),
    values = color.scheme.3
    ) +
  labs(
    x = p.label.nm,
    y = p.label.norm.abs
  ) +
  theme(legend.position = "bottom")
```


### All time points 

Note: Time in y-axis is not proportionally represented to real time 

```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.full.spectrum %>% 
  ggplot(aes(x = nm, y = as.factor(time_h), height = norm_abs, colour = induction, fill = induction, group = interaction(time_h, induction))) +
  geom_ridgeline(alpha = 0.05)  +
    facet_grid(location~strain) +
  labs(
    x = p.label.nm,
    y = p.label.time.h
  ) +
  theme(legend.position = "bottom")
```

```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}

bind_rows(
  df.full.spectrum %>% 
  filter(induction == "Time 0") %>% 
  mutate(
    induction = "+"
  ),
  df.full.spectrum %>% 
  filter(induction == "Time 0") %>% 
  mutate(
    induction = "-"
  ) 
) %>% 
  bind_rows(
    df.full.spectrum %>% 
  filter(induction != "Time 0")    
  ) %>% 
  ggplot(aes(x = nm, y = as.factor(time_h), height = norm_abs, colour = location, fill = location, group = interaction(time_h, location))) +
  geom_ridgeline(alpha = 0.05)  +
  facet_grid(induction~strain) +
  labs(y = "Time (h)") +
  scale_colour_manual(
    breaks = names(color.scheme.3),
    values = color.scheme.3,
    
    ) +
  scale_fill_manual(
    breaks = names(color.scheme.3),
    values = color.scheme.3
    ) +
  labs(
    x = p.label.nm,
    y = p.label.norm.abs,
    colour = ""
  ) +
  guides(fill = "none") +
  theme(legend.position = "bottom")
```