---
title: "Interlab study"
subtitle: "Test run"
author: "Hugo Pineda Hernandez"
date: "`r strftime(Sys.Date(), '%B %d %Y')`"
output: 
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning=FALSE, message = FALSE
  
)

library("tidyverse")
library(ggthemes)
library(cowplot)
library(ggridges)
theme_set(theme_light(14))
```

```{r load_data, include=FALSE}
# Load project settings
source(here::here("settings.R"))
# load functions
source(file.path(data.raw.dir, "experiments.R"))

df.pr.od.raw <- load_plate_reader_od()

df.pr.fl.raw <- load_plate_reader_fl()

df.spectrophotometer <- load_spectrophotometer() 

df.full.spectrum.raw <- load_full_spectrum()
```

```{r}

df.pr.od.blank <- df.pr.od.raw %>% 
  filter(!is.na(OD_730), sample_type == "blank") %>% 
  group_by(experiment_date, location, time_h) %>% 
  summarise(
    blank_OD_730 = mean(OD_730)
  ) %>% 
  ungroup()
  
df.pr.od <- df.pr.od.raw %>% 
  filter(!is.na(OD_730), sample_type == "sample") %>% 
  left_join(df.pr.od.blank, by = c("experiment_date", "location", "time_h")) %>% 
  mutate(
    # correct with blank
    OD_730_bc = OD_730 - blank_OD_730
  ) 
  
```

```{r}

df.pr.fl.blank <- df.pr.fl.raw %>% 
  filter(sample_type == "blank") %>% 
  group_by(experiment_date, location, time_h) %>% 
  summarise(
    blank_fl = mean(fl)
  ) %>% 
  ungroup()
  

df.pr.fl <- df.pr.fl.raw %>% 
  filter(sample_type == "sample") %>% 
   left_join(df.pr.fl.blank,
              by = c("experiment_date", "location", "time_h")) %>% 
  mutate(
    # correct with blank
    fl_bc = fl - blank_fl
  ) 
```


```{r}
df.pr <- full_join(
  df.pr.od, 
  df.pr.fl,
  by = c("strain", "induction", "bio_replicate", "location", "tech_replicate", "experiment_date", "experiment_id", "time_h", "sample_type")
  )  
  #   mutate(
  #   fl_od = fl_bc / OD_730_bc
  # ) 
```



```{r}

color.scheme.4 <- c("Amsterdam" = "#D55E00", 
                  "Düsseldorf" = "#CC79A7",
                  "Jena" = "#009E73",
                  "Seville" = "#0072B2")

color.scheme.3 <- c("Amsterdam" = "#D55E00", 
                  "Jena" = "#009E73",
                  "Seville" = "#0072B2")


plot_summary_location_strain <- function(){
  list(
  stat_summary(fun.data = "mean_cl_boot", geom = "point", alpha = 0.6),
  stat_summary(fun.data = "mean_sdl", geom = "errorbar", alpha = 0.6),
  stat_summary(fun.data = "mean_cl_boot", geom = "line", alpha = 0.6),
  facet_grid(location~strain, scales = "free_y"),
    scale_colour_colorblind(),
  theme(legend.position = "bottom")
  )
}

plot_summary_ind_strain <- function(colors.scheme){
  list(
  stat_summary(fun = "mean", geom = "point", alpha = 0.6),
  # stat_summary(fun.data = "mean_sdl", geom = "errorbar", alpha = 0.6),
  stat_summary(fun = "mean", geom = "line", alpha = 0.6),
  facet_grid(induction~strain),
  expand_limits(y = 0),
  scale_colour_manual(
    breaks = names(colors.scheme),
    values = colors.scheme
    ),
  theme(legend.position = "bottom")
  )
}
```


# Plate reader data

## Plate reader raw data

### OD<sub>730</sub>

```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr.od %>% 
  ggplot(aes(x = time_h, y = OD_730, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = p.label.od.730
  )
```

```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr.od %>% 
  ggplot(aes(x = time_h, y = OD_730, colour = location)) +
 plot_summary_ind_strain(colors.scheme = color.scheme.3) + 
  labs(
    x = p.label.time.h,
    y = p.label.od.730,
    colour = ""
  ) 
```

### Fluorescence

```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr.fl %>% 
  ggplot(aes(x = time_h, y = fl, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = "Raw flourescence"
  ) 
```


```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr.fl %>% 
  ggplot(aes(x = time_h, y = fl, colour = location)) +
   plot_summary_ind_strain(colors.scheme = color.scheme.4) + 
  labs(
    x = p.label.time.h,
    y = "Raw flourescence",
    colour = ""
  ) 
```



## Plate reader data normalization

* Step 1, background OD<sub>730</sub> correction:
  - Calculate average OD<sub>730</sub> of blank wells for each timepoint
  - Subtract blank average OD<sub>730</sub> from the rest of the wells for each timepoint

* Step 2, EVC normalization:
  - For each timepoint, calculate the average raw fluorescence / OD<sub>730</sub> of ALL the EVC wells
  - For each timepoint, calculate raw fluorescence / OD<sub>730</sub> for each well
  - For each timepoint, subtract the average EVC fluorescence / OD<sub>730</sub> from the fluorescence / OD<sub>730</sub> of each well
  
This step corrects for autofluorescence and background fluorescence at the same time and it also normalizes for the biomass.   

* Step 3, summarize:
  - Take the average of the background corrected fluorescence / OD<sub>730</sub> caculated in the previous step, for each experiment, strain, condition and timepoint.
  
* Step 4, 0 to 1 normalization:
  - For each experiment, normalize the average ratios calculated previously from 0 to 1 by using the min max method:
    $\frac{\frac{F}{OD_{730}} - min(\frac{F}{OD_{730}})}{ max(\frac{F}{OD_{730}}) -  min(\frac{F}{OD_{730}})}$

This step allows to compare the ratios between labs.
  
```{r}

df.pr.evc <- df.pr %>%
  filter(strain == "EVC", !is.na(OD_730_bc)) %>%
  group_by(location, experiment_date, time_h) %>%
  mutate(
    fl_od_evc = fl / OD_730_bc
  ) %>%
  summarise(
    fl_od_evc = mean(fl_od_evc)
  ) %>%
  ungroup()


df.pr.norm <- df.pr %>%
  filter(!is.na(OD_730_bc)) %>%
  left_join(df.pr.evc, by = c("location", "experiment_date", "time_h")) %>%
  group_by(location, experiment_date, strain, induction, tech_replicate, time_h) %>%
  mutate(
    fl_od = fl / OD_730_bc,
    fl_od_norm = fl_od - fl_od_evc
  ) %>% 
  group_by(location, experiment_date, strain, induction, time_h) %>%
  summarise(
    fl_od_norm = mean(fl_od_norm)
  ) %>%
  group_by(location, experiment_date) %>%
  mutate(
    fl_od_norm_0_to_1 = (fl_od_norm - min(fl_od_norm)) / (max(fl_od_norm) - min(fl_od_norm))
  ) %>%
  ungroup()

```
   
   
### Results after step 3

Mind the different scales in Y-axis.

```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr.norm %>% 
  ggplot(aes(x = time_h, y = fl_od_norm, colour = induction)) +
  plot_summary_location_strain() +
  labs(
    x = p.label.time.h,
    y = bquote("Background corrected fluorescence"%.%"OD"[730]^-1)
  ) 
```

```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr.norm %>% 
  ggplot(aes(x = time_h, y = fl_od_norm, colour = location)) +
  plot_summary_ind_strain(colors.scheme = color.scheme.3) +
  labs(
    x = p.label.time.h,
    y = bquote("Background corrected fluorescence"%.%"OD"[730]^-1)
  ) 
```

### Results after step 4

```{r, fig.width=11.11, fig.height=7, out.width="100%"}
df.pr.norm %>% 
  ggplot(aes(x = time_h, y = fl_od_norm_0_to_1, colour = induction)) +
  plot_summary_location_strain() +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = p.label.time.h,
    y = bquote("Relative fluorescence"%.%"OD"[730]^-1~"(%)")
  ) 
```

```{r, fig.width=11.11, fig.height=6.25, out.width="100%"}
df.pr.norm %>% 
  ggplot(aes(x = time_h, y = fl_od_norm_0_to_1, colour = location)) +
  plot_summary_ind_strain(colors.scheme = color.scheme.3) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = p.label.time.h,
    y = bquote("Relative fluorescence"%.%"OD"[730]^-1~"(%)")
  ) 
```
